---
title: "Bank Credit's Customer Dashboard"
author: "MinhChau"
output: 
  html_document: default
---

```{css, echo=FALSE}

.rpivotTable{ overflow-x: scroll; }
.rpivotTable{ overflow: auto; }

```

```{r load-libraries, echo=FALSE}
#install.packages - only need to run this code once to install the package
# load required packages
library(dplyr)
library(tidyverse) #need to call the library before you use the packages
library(rpivotTable)
library(knitr)
library(readxl) # this is required for the read_excel function
Bank_Credit_Data <- read_excel("/Users/minhchau/Downloads/Bank Credit Risk Data.xlsx", sheet="Base Data", skip=2)
BD <- Bank_Credit_Data
head(BD)
```

### 1. Customer Profile Dashboard for Credit Risk Analysis 

- i.View the frequency distributions for each of the following customer demographic variables in a chart and table (i.e. one table and chart per variable): `Credit Risk` and `Total`. `Total` is the sum of `Checking` and `Savings`. (hint: You will need to create the variable `Total` in the dataframe). 
- ii. View the relationship between `Total` and `Months Employed` in one chart, and the relationship between `Total` and `Age` in another chart.
- iii. Provide a description of any interesting patterns observed from the charts. (You may type your answer in the space below)
- iv. Frequency distributions for  `Housing`, `Job` and `Months Employed` on your own. 

```{r q1.(a), echo=TRUE}
BD$Total <- BD$Checking+BD$Savings
BD <- BD %>%
  mutate(Total = Checking+Savings)

crFreq <- BD%>%count(`Credit Risk`)
kable(crFreq, caption = "Frequency of Bank Customers by Credit Risk")
slice.cr <- crFreq$n
cr.piepercent <- 100*round(crFreq$n/sum(crFreq$n),2)
label<-crFreq$`Credit Risk`
label<-paste(label, ",", sep="")
label<-paste(label,cr.piepercent)
label<-paste(label, "%", sep="")
pie(slice.cr,
    labels=label,
    col=c("blue", "cyan"),
    radius=1,
    main="Customer Credit Risk")
h.tot <- hist(BD$Total,
              main="Histogram of customer total amount balance",
              xlab="Total account balance",
              ylab="No. of Customers",
              col=c("darkorange"),
              ylim=c(0,400),
              labels=TRUE)
ab.Group <- cut(BD$Total, h.tot$breaks, include.lowest = TRUE, dig.lab=5)
t.emp<- table(ab.Group)
kable(t.emp, caption = "Frequency distribution by Total Account Balance")
plot(BD$Total,
     BD$`Months Employed`,
     main="Scatterplot of months memployed  to total",
     ylab="Months employed", xlab="Total")
plot(BD$Total, BD$Age, main="Scatterplot of Age  to Total",ylab="Age", xlab="Total")

JobFreq <- BD%>% count(Job)
kable(JobFreq, caption="Frequency of Bank Customers by Job")

slice.job <- JobFreq$n
job.piepercent <- 100*round(JobFreq$n/sum(JobFreq$n),2)
label <- JobFreq$Job
label <- paste(label, ",", sep="")
label <- paste(label, job.piepercent)
label <- paste(label, "%", sep="")
pie(slice.job,
    labels = label,
    col=c("blue", "cyan", "dodgerblue", "deepskyblue"),
    radius=1,
    main="Customer Job")
```

### 2. Customers' Demographics 
- i. see the appropriate chart and table to compare frequency of customers by `Credit Risk` and `Job`.  
- ii. see a description of any interesting patterns observed from the charts.

```{r q1.(b), echo=TRUE}

BDb1 <- BD %>% group_by(`Credit Risk`, Job)%>% tally()
BDb1.spread <- BDb1 %>% spread(key = Job, value=n)
kable(BDb1.spread, caption = "Contrigency table for Credit risk and Job")
barmatrix.BDb1 <- as.matrix(BDb1.spread[,c(2:5)])
barcol <-c("blue", "gray")
barplot(barmatrix.BDb1,
        col=barcol,
        main="Frequency of Customer by Credit Risk and Job",
        ylab="No.og customers",
        xlab="Job")
legend("topright",
       cex=0.6,
       fill=barcol,
       legend=BDb1.spread$`Credit Risk`,
       title="Credit Risk")
barplot(barmatrix.BDb1,
        col=barcol,
        main="Frequency of Customer by Credit risk and Job",
        ylab="No.of Customers",
        xlab="Job",
        beside=TRUE)
legend("topright", cex=0.6, fill=barcol, legend=BDb1.spread$`Credit Risk`, title = "Credit Risk")

```

### 3. Credit Risk Analysis 

- i. View the the `Loan Purpose` of customers with "High" levels of `Credit Risk`. Visualize the frequency distribution of `Loan Purpose` for "High" `Credit Risk` customers
- ii. The most and least common `Loan Purpose` types among "High" `Credit Risk` customers

```{r q1.(c), echo=TRUE}
LoanHRFreq<- BD%>%
  filter (`Credit Risk` == "High") %>%
  count(`Loan Purpose`)
kable(LoanHRFreq, caption = "Frequency Distribution for Loan Purpose for High CR Customers")
LoanHRbar <- LoanHRFreq$n
bploanHR <- barplot(LoanHRbar, names.arg = LoanHRFreq$`Loan Purpose`,
        col="blue",
        main="Frequency of Loan Purpose for High CR Customers",
        cex.names = 0.8,
        xlim=c(0,80),
        xlab="No. of Loans",
        horiz=TRUE,
        las=1)
text(x=LoanHRbar, y=bploanHR, col="black", LoanHRFreq$n, cex=0.8, pos=4)
LoanMax <- LoanHRFreq%>% slice_max(n, n=1)
LoanMin <- LoanHRFreq%>% slice_min(n, n=1)
rbind(LoanMax, LoanMin)

```

Q1.(d) Customer Account Balance Pareto Analyses
Show the number and percentage of customers that contribute most, amounting to 80% of the total account balanced with the bank. 

```{r q1.(d), echo=TRUE}

BD.tot <- BD %>%
  select(Total)%>%
  arrange(desc(Total))
BD.tot$Percentage <-BD.tot$Total/sum(BD.tot$Total)
BD.tot$Cumulative <- cumsum(BD.tot$Percentage)
BD.tot$Cumulative.cust<-as.numeric(rownames(BD))/nrow(BD)
which(BD.tot$Cumulative>0.8)[1]
(which(BD.tot$Cumulative>0.8)[1])/nrow(BD)
```
